---
title: "Student Centers of Interest and Rent Prices - R Notebook"
output:
  html_document:
    toc: true
    toc_depth: '2'
    df_print: paged
  html_notebook:
    toc: true
    toc_depth: 2
---

# How to run this file

#Put this `.Rmd` file in the same folder as the CSV file. Then open this file in RStudio and run each gray R chunk using the green **Run** button.  
#Do **not** copy the lines that start with ``` into the Console.

# 1. Packages

```{r , message=FALSE, warning=FALSE}
# If a package is missing, install it automatically
required_packages <- c(
  "dplyr", "readr", "tidyr", "ggplot2", "corrplot",
  "randomForest", "caret", "car", "lmtest", "tseries", "xgboost"
)

install_if_missing <- function(pkg) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg)
  }
}

invisible(lapply(required_packages, install_if_missing))

library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(corrplot)
library(randomForest)
library(caret)
library(car)
library(lmtest)
library(tseries)
library(xgboost)
library(MASS) 
library(tibble)
library(fastDummies)
library(mgcv)

set.seed(42)
```

# 2. Load the CSV file

```{r , message=FALSE, warning=FALSE}
# Change only this file name if your CSV has a different name
INPUT_CSV <- "appartments_dedup_equal_query.csv"

# Load data
df <- read_csv(INPUT_CSV, show_col_types = FALSE)

# Show dimensions and first rows
cat("Rows:", nrow(df), "\n")
cat("Columns:", ncol(df), "\n")
head(df)
glimpse(df)
```

# 3. Basic cleaning

```{r , message=FALSE, warning=FALSE}
# Keep only positive prices and rows with core apartment information
df <- df %>%
  filter(price > 0) %>%
  filter(price > 1500) %>%
  filter(price < 8000) %>%
  drop_na(price, bedrooms, squareFeet)

# Convert Boolean columns to 0/1 when they exist
bool_cols <- c("hasBalcony", "hasMamad", "hasShelter", "isFlatshare", "isBroker")

for (col in bool_cols) {
  if (col %in% names(df)) {
    df[[col]][is.na(df[[col]])] <- FALSE
    df[[col]] <- as.integer(as.logical(df[[col]]))
  }
}

# Keep rows with calculated walking distance to BGU gate
if ("walking_distance_bgu_gate_m" %in% names(df)) {
  df <- df %>% drop_na(walking_distance_bgu_gate_m)
}

# Make sure student hub distance exists
if (!("walking_distance_student_hub_m" %in% names(df))) {
  hub_cols <- c(
    "walking_distance_ringelblum_m",
    "walking_distance_block_center_m",
    "walking_distance_sophrim_park_m"
  )
  hub_cols <- hub_cols[hub_cols %in% names(df)]
  if (length(hub_cols) > 0) {
    df$walking_distance_student_hub_m <- apply(df[, hub_cols], 1, min, na.rm = TRUE)
  }
}

# Reset row numbers
row.names(df) <- NULL

cat("Total apartments ready for analysis:", nrow(df), "\n")
df %>% count(neighborhood, sort = TRUE)

```

# 4. Correlation matrix (Log Transformed Distances)

```{r , message=FALSE, warning=FALSE, fig.width=12, fig.height=10}
# 1. Define baseline numerical features
base_features <- c("price", "bedrooms", "squareFeet", "hasBalcony", "hasMamad")
base_features <- base_features[base_features %in% names(df)]

# 2. Extract and filter walking distance columns
distance_cols <- grep("^walking_distance_", names(df), value = TRUE)

# 3. Process data and convert categorical neighborhood into dummy variables (One-Hot Encoding)
corr_df <- df %>%
  # Generate explicit 0/1 columns for each neighborhood level (e.g., neighborhood_Gimel, neighborhood_Shchuna D)
  fastDummies::dummy_cols(select_columns = "neighborhood", remove_selected_columns = TRUE) %>%
  # Keep core numerical features, newly created neighborhood dummies, and distance columns
  dplyr::select(all_of(base_features), starts_with("neighborhood_"), all_of(distance_cols)) %>%
  mutate(across(everything(), as.numeric)) %>%
  mutate(across(all_of(distance_cols), ~ log(. + 1)))

# 4. Clean column names for professional and scannable visual labeling
names(corr_df) <- gsub("walking_distance_", "log_", names(corr_df))
names(corr_df) <- gsub("neighborhood_", "hood_", names(corr_df)) # Shorten prefix for plot alignment

# 5. Compute the correlation matrix
corr_matrix <- cor(corr_df, use = "pairwise.complete.obs")

# 6. Generate the final styled heatmap
corrplot(
  corr_matrix,
  method = "color",
  addCoef.col = "black",
  tl.col = "black",
  tl.srt = 45,
  col = colorRampPalette(c("blue", "white", "red"))(200),
  title = "Correlation Matrix - Rent Prices, Hoods & Log Student Centers",
  mar = c(0, 0, 2, 0),
  number.cex = 0.6 # Slightly downsized to fit all new dummy columns perfectly
)
```

# 5. Association Matrix: Neighborhood as a Single Feature

```{r , message=FALSE, warning=FALSE, fig.width=12, fig.height=10}
# 1. Define variables to test against neighborhood
target_features <- c("price", "bedrooms", "squareFeet", "hasBalcony", "hasMamad")
target_features <- target_features[target_features %in% names(df)]
distance_cols <- grep("^walking_distance_", names(df), value = TRUE)
all_vars <- c(target_features, distance_cols)

# 2. Calculate Eta-squared for neighborhood against each variable using ANOVA
eta_results <- sapply(all_vars, function(var) {
  # Clean names and apply log to distances just like before
  if (var %in% distance_cols) {
    y_val <- log(as.numeric(df[[var]]) + 1)
  } else {
    y_val <- as.numeric(df[[var]])
  }
  
  # Run a simple ANOVA model: Variable ~ Neighborhood
  fit <- lm(y_val ~ as.factor(df$neighborhood))
  
  # Extract Sum of Squares to calculate Eta-squared manually
  anova_summary <- anova(fit)
  ss_neighborhood <- anova_summary["as.factor(df$neighborhood)", "Sum Sq"]
  ss_total <- sum(anova_summary["Sum Sq"])
  
  return(ss_neighborhood / ss_total)
})

# 3. Construct a flat data frame for ggplot visualization
heatmap_data <- data.frame(
  Variable = names(eta_results),
  Neighborhood = "neighborhood",
  Association = as.numeric(eta_results)
)

# Clean up feature labels for a professional look
heatmap_data$Variable <- gsub("walking_distance_", "log_", heatmap_data$Variable)

# 4. Generate the styled single-column heatmap
ggplot(heatmap_data, aes(x = Neighborhood, y = Variable, fill = Association)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(Association, 2)), color = "black", fontface = "bold") +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0.5, limits = c(0,1)) +
  labs(
    title = "Association Matrix: Neighborhood as a Single Feature",
    subtitle = "Values represent Eta-squared (Variance Explained by Hood)",
    x = "",
    y = "Features"
  ) +
  theme_minimal()
```

# 6. Price distribution

```{r , message=FALSE, warning=FALSE, fig.width=10, fig.height=6}
mean_price <- mean((df$price), na.rm = TRUE)
median_price <- median((df$price), na.rm = TRUE)

ggplot(df, aes(x = (price))) +
  geom_histogram(aes(y = after_stat(density)), bins = 30, fill = "#008080", color = "black", alpha = 0.7) +
  geom_density(color = "darkblue", linewidth = 1) +
  geom_vline(xintercept = mean_price, color = "red", linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = median_price, color = "orange", linewidth = 1) +
  labs(
    title = "Distribution of Rental Prices",
    subtitle = paste0("Mean  = ", round(mean_price, 2), ", Median  = ", round(median_price, 2)),
    x = "Price",
    y = "Density"
  ) +
  theme_minimal()
```

# 7. Log Price distribution

```{r , message=FALSE, warning=FALSE, fig.width=10, fig.height=6}
mean_log_price <- mean(log(df$price), na.rm = TRUE)
median_log_price <- median(log(df$price), na.rm = TRUE)

ggplot(df, aes(x = log(price))) +
  geom_histogram(aes(y = after_stat(density)), bins = 30, fill = "#008080", color = "black", alpha = 0.7) +
  geom_density(color = "darkblue", linewidth = 1) +
  geom_vline(xintercept = mean_log_price, color = "red", linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = median_log_price, color = "orange", linewidth = 1) +
  labs(
    title = "Distribution of Log Rental Prices",
    subtitle = paste0("Mean (log) = ", round(mean_log_price, 2), ", Median (log) = ", round(median_log_price, 2)),
    x = "Log Price",
    y = "Density"
  ) +
  theme_minimal()
```


# 8. Rental prices by neighborhood

```{r , message=FALSE, warning=FALSE, fig.width=10, fig.height=6}
ggplot(df, aes(x = neighborhood, y = log(price), fill = neighborhood)) +
  geom_boxplot() +
  labs(
    title = "Rental Prices by Neighborhood",
    x = "Neighborhood",
    y = "log price (NIS)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

# 9. Simple regressions: log price versus log walking distance

```{r , message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
distance_cols <- grep("^walking_distance_", names(df), value = TRUE)

plot_df <- data.frame()

for (col in distance_cols) {
  log_price_vals <- log(df$price)
  log_dist_vals <- log(df[[col]] + 1)
  
  price_q <- quantile(log_price_vals, probs = c(0.05, 0.95), na.rm = TRUE)
  dist_q <- quantile(log_dist_vals, probs = c(0.05, 0.95), na.rm = TRUE)

  temp <- data.frame(
    log_price = log_price_vals,
    log_distance = log_dist_vals,
    orig_dist = df[[col]]
  ) %>%
    filter(
      log_price >= price_q[1], log_price <= price_q[2],
      log_distance >= dist_q[1], log_distance <= dist_q[2]
    ) %>%
    drop_na()

  if (nrow(temp) >= 10) {
    temp$center <- gsub("walking_distance_", "", col)
    temp$center <- gsub("_m", "", temp$center)
    temp$center <- gsub("_", " ", temp$center)
    plot_df <- bind_rows(plot_df, temp %>% dplyr::select(log_price, log_distance, center))
  }
}

ggplot(plot_df, aes(x = log_distance, y = log_price)) +
  geom_point(alpha = 0.55) +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  facet_wrap(~ center, scales = "free_x") +
  labs(
    title = "Log Price vs Log Walking Distance (+1) to Centers of Interest",
    x = "Log Walking Distance (meters + 1)",
    y = "Log Rental Price"
  ) +
  theme_minimal()
```

# 10. Simple regressions summary table

```{r , message=FALSE, warning=FALSE}
simple_results <- data.frame()

for (col in distance_cols) {
  log_price_vals <- log(df$price)
  log_dist_vals <- log(df[[col]] + 1)
  
  price_q <- quantile(log_price_vals, probs = c(0.05, 0.95), na.rm = TRUE)
  dist_q <- quantile(log_dist_vals, probs = c(0.05, 0.95), na.rm = TRUE)

  temp <- data.frame(
    log_price = log_price_vals,
    log_distance = log_dist_vals
  ) %>%
    filter(
      log_price >= price_q[1], log_price <= price_q[2],
      log_distance >= dist_q[1], log_distance <= dist_q[2]
    ) %>%
    drop_na()

  if (nrow(temp) >= 10) {
    fit <- lm(log_price ~ log_distance, data = temp)
    sm <- summary(fit)
    coef_row <- sm$coefficients["log_distance", ]

    simple_results <- bind_rows(simple_results, data.frame(
      Center_of_Interest = tools::toTitleCase(gsub("_", " ", gsub("_m", "", gsub("walking_distance_", "", col)))),
      Variable = col,
      Coef = coef_row["Estimate"],
      Std_Error = coef_row["Std. Error"],
      t_value = coef_row["t value"],
      p_value = coef_row["Pr(>|t|)"],
      Direction = ifelse(coef_row["Estimate"] > 0, "Positive", "Negative"),
      R_squared = sm$r.squared,
      Adj_R_squared = sm$adj.r.squared,
      N = nrow(temp)
    ))
  }
}

simple_results <- simple_results %>%
  mutate(
    Significance = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01 ~ "**",
      p_value < 0.05 ~ "*",
      p_value < 0.1 ~ ".",
      TRUE ~ "Not significant"
    )
  ) %>%
  arrange(p_value)

simple_results
```

# 11. Feature engineering for full models

```{r , message=FALSE, warning=FALSE}
engineer_features <- function(data) {
  data <- data %>% mutate(squareFeet = ifelse(squareFeet == 0, NA, squareFeet))

  data <- data %>%
    mutate(
      roommates = ifelse(bedrooms > 1, bedrooms - 1, bedrooms),
      roommates = ifelse(roommates <= 0 | is.na(roommates), 1, roommates),
      price_per_roommate = price / roommates,
      log_price = log(price), 
      sqft_per_bedroom = squareFeet / ifelse(bedrooms == 0, 0.5, bedrooms)
    )

  # Spaciousness z-score within neighborhood
  data <- data %>%
    group_by(neighborhood) %>%
    mutate(spaciousness_zscore = as.numeric(scale(sqft_per_bedroom))) %>%
    ungroup() %>%
    mutate(spaciousness_zscore = ifelse(is.na(spaciousness_zscore), 0, spaciousness_zscore))

  distance_cols_local <- grep("^walking_distance_", names(data), value = TRUE)
  for (col in distance_cols_local) {
    data[[paste0("log_", col)]] <- log(data[[col]] + 1)
  }

  bool_cols_local <- c("hasBalcony", "hasMamad")
  for (col in bool_cols_local) {
    if (col %in% names(data)) {
      data[[col]] <- as.numeric(data[[col]])
    }
  }

  return(data)
}

df_engineered <- engineer_features(df)
head(df_engineered %>% dplyr::select(price, log_price, roommates, spaciousness_zscore, grep("^log_walking_", names(df_engineered), value = TRUE)[1:2]))
```

# 12. Full OLS model with Outlier Removal (Log-Log Model)

```{r , message=FALSE, warning=FALSE}

selected_distances <- c(
  "log_walking_distance_bgu_gate_m",
  "log_walking_distance_ringelblum_m",
  "log_walking_distance_sophrim_park_m",
  "log_walking_distance_metzadah_m",
  "log_walking_distance_block_center_m",
  "log_walking_distance_nearest_bar_m",
  "log_walking_distance_nearest_cafe_m",
  "log_walking_distance_nearest_restaurant_m"
)

selected_distances <- selected_distances[selected_distances %in% names(df_engineered)]

base_features <- c("neighborhood", "roommates", "spaciousness_zscore", "hasBalcony", "hasMamad")
base_features <- base_features[base_features %in% names(df_engineered)]

required_cols <- c("log_price", base_features, selected_distances)
model_df <- df_engineered %>% dplyr::select(all_of(required_cols)) %>% drop_na()
model_df$neighborhood <- as.factor(model_df$neighborhood)

ols_formula <- as.formula(
  paste("log_price ~", paste(c("neighborhood", setdiff(base_features, "neighborhood"), selected_distances), collapse = " + "))
)

temp_model <- lm(ols_formula, data = model_df)

cooks_d <- cooks.distance(temp_model)
influential_obs <- which(cooks_d > (4 / nrow(model_df)))

if (length(influential_obs) > 0) {
  model_df <- model_df[-influential_obs, ]
}

final_model <- lm(ols_formula, data = model_df)

cat("--- OLS Model Summary (Cleaned Log-Log Model) ---\n")
cat("Removed", length(influential_obs), "influential observations.\n\n")
summary(final_model)

cat("\n--- Variance Inflation Factor (VIF) ---\n")
cat("Scores above 5 may indicate problematic multicollinearity.\n")
print(vif(final_model))
```

# 13. Full OLS diagnostics

```{r , message=FALSE, warning=FALSE, fig.width=14, fig.height=5}
residuals_ols <- resid(final_model)
fitted_ols <- fitted(final_model)

diagnostic_df <- data.frame(fitted = fitted_ols, residuals = residuals_ols)

# Residuals vs Fitted- Testing the linearity assumption
p1 <- ggplot(diagnostic_df, aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Residuals vs Fitted Values", x = "Fitted Values", y = "Residuals") +
  theme_minimal()

# Histogram of residuals - -Testing the assumption of normality of errors
p2 <- ggplot(diagnostic_df, aes(x = residuals)) +
  geom_histogram(bins = 30, fill = "gray70", color = "black") +
  labs(title = "Histogram of Residuals", x = "Residuals", y = "Frequency") +
  theme_minimal()

print(p1)
print(p2)

# Q-Q plot -Testing the assumption of normality of errors
qqnorm(residuals_ols, main = "Q-Q Plot of OLS Residuals")
qqline(residuals_ols, col = "red")

#BP-Testing the assumption of homoscedasticity — equality of variances 
cat("\n--- Breusch-Pagan Test ---\n")
print(bptest(final_model))

#Testing the assumption of normality of errors
cat("\n--- Jarque-Bera Test ---\n")
print(jarque.bera.test(residuals_ols))

#DW- Residual independence test
cat("\n--- Durbin-Watson Test ---\n")
print(dwtest(final_model))

# Formal Testing for the assumption of linearity
cat("\n--- Rainbow Test for Linearity ---\n")
print(lmtest::raintest(final_model))
```

# 14. Full OLS coefficient plot

```{r , message=FALSE, warning=FALSE, fig.width=12, fig.height=7}
coef_summary <- summary(final_model)$coefficients
coef_df <- data.frame(
  Feature = rownames(coef_summary),
  Coef = coef_summary[, "Estimate"],
  Std_Error = coef_summary[, "Std. Error"],
  p_value = coef_summary[, "Pr(>|t|)"]
) %>%
  filter(Feature != "(Intercept)") %>%
  mutate(Significant = p_value < 0.05) %>%
  arrange(Coef)

coef_df$Feature <- factor(coef_df$Feature, levels = coef_df$Feature)

ggplot(coef_df, aes(x = Coef, y = Feature, fill = Significant)) +
  geom_col() +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    title = "OLS Coefficients",
    x = "Coefficient",
    y = "Feature"
  ) +
  theme_minimal()
```

#15. Reduced OLS model 

```{r , message=FALSE, warning=FALSE, fig.width=12, fig.height=7}
# 1. Selection & Estimation (Multiple Linear Regression - OLS)
selected_features <- c(
  "roommates", 
  "spaciousness_zscore",
  "hasMamad",
  "hasBalcony",
  "log_walking_distance_bgu_gate_m",
  "log_walking_distance_sophrim_park_m",
  "log_walking_distance_nearest_restaurant_m",
  "log_walking_distance_nearest_bar_m",
  "log_walking_distance_block_center_m"
)

optimal_formula <- as.formula(paste("log_price ~", paste(selected_features, collapse = " + ")))

# Fit the reduced Ordinary Least Squares (OLS) model
reduced_ols_model <- lm(optimal_formula, data = model_df)
reduced_summary <- summary(reduced_ols_model)

# 2. P-Values & Significance Table
reduced_coefs <- reduced_summary$coefficients
reduced_results_clean <- data.frame(
  Feature = rownames(reduced_coefs),
  Estimate = as.numeric(reduced_coefs[, "Estimate"]),
  Std_Error = as.numeric(reduced_coefs[, "Std. Error"]),
  t_value = as.numeric(reduced_coefs[, "t value"]),
  p_value = round(as.numeric(reduced_coefs[, "Pr(>|t|)"]), 5),
  Signif = case_when(
    reduced_coefs[, "Pr(>|t|)"] < 0.001 ~ "***", 
    reduced_coefs[, "Pr(>|t|)"] < 0.01 ~ "**", 
    reduced_coefs[, "Pr(>|t|)"] < 0.05 ~ "*", 
    reduced_coefs[, "Pr(>|t|)"] < 0.1 ~ ".", 
    TRUE ~ " "
  ),
  stringsAsFactors = FALSE
)

cat("--- Coefficients Table ---\n")
print(reduced_results_clean, row.names = FALSE)

# 3. Model Performance Metrics
cat("\n--- Performance Metrics ---\n")
cat("MAE:                ", round(mean(abs(resid(reduced_ols_model))), 4), "\n")
cat("RMSE:               ", round(sqrt(mean(resid(reduced_ols_model)^2)), 4), "\n")
cat("R-squared:          ", round(reduced_summary$r.squared, 4), "\n")
cat("Adjusted R-squared: ", round(reduced_summary$adj.r.squared, 4), "\n")

# 4. Coefficients Plot
plot_data_reduced <- reduced_results_clean %>%
  filter(Feature != "(Intercept)") %>%
  mutate(Significant = ifelse(p_value < 0.05, "TRUE", "FALSE")) %>%
  arrange(Estimate)

plot_data_reduced$Feature <- factor(plot_data_reduced$Feature, levels = plot_data_reduced$Feature)

ggplot(plot_data_reduced, aes(x = Estimate, y = Feature, fill = Significant)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black", width = 0.7) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_fill_manual(values = c("TRUE" = "#008080", "FALSE" = "#FA8072")) +
  scale_y_discrete(labels = function(x) gsub("log_walking_distance_", "log_", x)) +
  theme_minimal()

# 5. Linear Regression Assumption Diagnostics
residuals_reduced <- resid(reduced_ols_model)

cat("\n--- Diagnostics: VIF (Multicollinearity) ---\n")
print(car::vif(reduced_ols_model))

cat("\n--- Diagnostics: Breusch-Pagan Test (Homoscedasticity) ---\n")
print(lmtest::bptest(reduced_ols_model))

cat("\n--- Diagnostics: Jarque-Bera Test (Normality of Errors) ---\n")
print(tseries::jarque.bera.test(residuals_reduced))

cat("\n--- Diagnostics: Durbin-Watson Test (Residual Independence) ---\n")
print(lmtest::dwtest(reduced_ols_model))

cat("\n--- Diagnostics: Rainbow Test (Linearity) ---\n")
print(lmtest::raintest(reduced_ols_model, fraction = 0.5))
```

#16. Quadratic OLS Regression Moddle (with ^2) 

```{r , message=FALSE, warning=FALSE, fig.width=12, fig.height=7}

# 1. Define formula with quadratic terms and fit the standard OLS linear regression model
quadratic_formula <- log_price ~ roommates + spaciousness_zscore + hasMamad + hasBalcony + 
  log_walking_distance_bgu_gate_m + I(log_walking_distance_bgu_gate_m^2) +
  log_walking_distance_sophrim_park_m +
  log_walking_distance_nearest_restaurant_m +
  log_walking_distance_nearest_bar_m + I(log_walking_distance_nearest_bar_m^2) +
  log_walking_distance_block_center_m

quadratic_ols_model <- lm(quadratic_formula, data = model_df)
ols_summary <- summary(quadratic_ols_model)

# 2. Extract Coefficients and model significance
ols_coefs <- ols_summary$coefficients
ols_results_clean <- data.frame(
  Feature = rownames(ols_coefs),
  Estimate = as.numeric(ols_coefs[, "Estimate"]),
  p_value = round(as.numeric(ols_coefs[, "Pr(>|t|)"]), 5),
  Signif = ifelse(ols_coefs[, "Pr(>|t|)"] < 0.001, "***",
           ifelse(ols_coefs[, "Pr(>|t|)"] < 0.01, "**",
           ifelse(ols_coefs[, "Pr(>|t|)"] < 0.05, "*",
           ifelse(ols_coefs[, "Pr(>|t|)"] < 0.1, ".", " ")))),
  stringsAsFactors = FALSE
)

cat("--- Coefficients Table ---\n")
print(ols_results_clean, row.names = FALSE)

# 3. Calculate model performance metrics
ols_res <- resid(quadratic_ols_model)
cat("\n--- Performance Metrics ---\n")
cat("MAE:                ", round(mean(abs(ols_res)), 4), "\n")
cat("RMSE:               ", round(sqrt(mean(ols_res^2)), 4), "\n")
cat("Robust R-squared:   ", round(ols_summary$r.squared, 4), "\n")
cat("Adjusted R-squared: ", round(ols_summary$adj.r.squared, 4), "\n")

# 4. Generate the coefficients visualization plot
plot_data_quad <- ols_results_clean[ols_results_clean$Feature != "(Intercept)", ]
plot_data_quad$Significant <- ifelse(plot_data_quad$p_value < 0.05, "TRUE", "FALSE")
plot_data_quad <- plot_data_quad[order(plot_data_quad$Estimate), ]
plot_data_quad$Feature <- factor(plot_data_quad$Feature, levels = plot_data_quad$Feature)

ggplot(plot_data_quad, aes(x = Estimate, y = Feature, fill = Significant)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black", width = 0.7) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1) +
  scale_fill_manual(values = c("TRUE" = "#008080", "FALSE" = "#FA8072")) +
  scale_y_discrete(labels = function(x) gsub("log_walking_distance_", "log_", x)) +
  labs(
    title = "OLS Regression Coefficients (With Quadratic Terms)",
    subtitle = "Estimated effect including squared terms for non-linear detection",
    x = "Coefficient Estimate",
    y = "Explanatory Variable",
    fill = "Significant"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 22, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 16, hjust = 0.5),
    axis.title.x = element_text(size = 16, face = "bold"),
    axis.title.y = element_text(size = 16, face = "bold"),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14)
  )

# 5. Linear Regression Assumption Diagnostics
residuals_reduced <- resid(quadratic_ols_model)

cat("\n--- Diagnostics: VIF (Multicollinearity) ---\n")
print(car::vif(quadratic_ols_model))

cat("\n--- Diagnostics: Breusch-Pagan Test (Homoscedasticity) ---\n")
print(lmtest::bptest(quadratic_ols_model))

cat("\n--- Diagnostics: Jarque-Bera Test (Normality of Errors) ---\n")
print(tseries::jarque.bera.test(residuals_reduced))

cat("\n--- Diagnostics: Durbin-Watson Test (Residual Independence) ---\n")
print(lmtest::dwtest(quadratic_ols_model))

cat("\n--- Diagnostics: Rainbow Test (Linearity) ---\n")
print(lmtest::raintest(quadratic_ols_model, fraction = 0.5))
```



#17. GAM model Performance & Goodness of Fit

```{r gam_model, message=FALSE, warning=FALSE, fig.width=12, fig.height=7}
# 1. Model Definition & Estimation (Generalized Additive Model - GAM)
gam_model <- gam(
  log_price ~ roommates + spaciousness_zscore + hasMamad + hasBalcony + 
    s(log_walking_distance_bgu_gate_m) + 
    s(log_walking_distance_sophrim_park_m) + 
    s(log_walking_distance_nearest_restaurant_m) + 
    s(log_walking_distance_nearest_bar_m),
  data = model_df, 
  method = "REML"
)

gam_summary <- summary(gam_model)

# 2. Results Extraction & Variable Significance Table
# Extract parametric (linear) coefficients
parametric_coefs <- gam_summary$p.table
parametric_df <- data.frame(
  Feature = rownames(parametric_coefs),
  Estimate = as.numeric(parametric_coefs[, "Estimate"]),
  p_value = as.numeric(parametric_coefs[, "Pr(>|t|)"]),
  Type = "Linear Asset Feature"
)

# Extract smooth (non-linear) distance features
# Note: EDF (Effective Degrees of Freedom) reflects non-linear complexity. EDF > 1 implies non-linearity.
smooth_coefs <- gam_summary$s.table
smooth_df <- data.frame(
  Feature = rownames(smooth_coefs),
  Estimate = as.numeric(smooth_coefs[, "edf"]), 
  p_value = as.numeric(smooth_coefs[, "p-value"]),
  Type = "Non-linear Distance Feature"
)

# Combine datasets for visualization and reporting
plot_data_gam <- rbind(parametric_df, smooth_df) %>%
  filter(Feature != "(Intercept)") %>%
  mutate(
    Significant = ifelse(p_value < 0.05, "TRUE", "FALSE"),
    # Clean feature names for clean plotting axis labels
    Feature = gsub("s\\(|\\)", "", Feature),
    Feature = gsub("log_walking_distance_", "log_", Feature)
  ) %>%
  arrange(Estimate)

plot_data_gam$Feature <- factor(plot_data_gam$Feature, levels = plot_data_gam$Feature)

cat("--- Coefficients & Variable Significance Table ---\n")
print(plot_data_gam %>% dplyr::select(Feature, p_value, Significant, Type), row.names = FALSE)

# 3. Model Performance Metrics
actual_y <- model_df$log_price
gam_preds <- as.numeric(predict(gam_model))
gam_res <- actual_y - gam_preds

cat("\n--- Model Performance Metrics ---\n")
cat("Deviance Explained (GAM R-squared): ", round(gam_summary$dev.expl * 100, 2), "%\n")
cat("MAE (Mean Absolute Error in Log):   ", round(mean(abs(gam_res)), 4), "\n")
cat("RMSE (Root Mean Squared Error):     ", round(sqrt(mean(gam_res^2)), 4), "\n")

# 4. Assumption Diagnostics
cat("\n--- Assumption Diagnostics ---\n")

# Assumption 1: Check if the basis size (k) chosen for smooth terms is sufficient
k_check <- k.check(gam_model)
cat("[Assumption 1 - Smoothness Flexibility (Basis Size)]:\n")
cat("If the p-value is > 0.05, the chosen basis size is sufficient and the assumption holds:\n")
print(k_check[, c("k'", "edf", "p-value")])

# Assumption 2: Multicollinearity check using a parallel standard OLS model
cat("\n[Assumption 2 - Multicollinearity Check (VIF)]:\n")
vif_ols <- lm(log_price ~ roommates + spaciousness_zscore + hasMamad + hasBalcony + 
                log_walking_distance_bgu_gate_m + log_walking_distance_sophrim_park_m + 
                log_walking_distance_nearest_restaurant_m + log_walking_distance_nearest_bar_m + 
                log_walking_distance_block_center_m, data = model_df)
print(car::vif(vif_ols))

# 5. Feature Importance Plot
ggplot(plot_data_gam, aes(x = Estimate, y = Feature, fill = Significant)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black", width = 0.7) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1) +
  scale_fill_manual(values = c("TRUE" = "#008080", "FALSE" = "#FA8072")) +
  labs(
    title = "GAM Feature Importance & Significance",
    subtitle = "Linear effects (Asset features) vs. Non-linear flexibility (Distance features)",
    x = "Effect Size (Estimate for Linear / EDF value for Smooth)",
    y = "Explanatory Variable",
    fill = "Is Significant?"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 11, hjust = 0.5),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10)
  )

# 6. Plotting Non-Linear Partial Spline Effects
# Visualizes the exact shapes of the non-linear relationship for each distance feature
plot(gam_model, pages = 1, shade = TRUE, main = "Partial Spline Effects of Distance Features")
```

# 18. Random Forest model

```{r , message=FALSE, warning=FALSE}
rf_features <- c(
  "neighborhood",
  "roommates",
  "spaciousness_zscore",
  "hasBalcony",
  "hasMamad",
  selected_distances
)

rf_df <- df_engineered %>%
  dplyr::select(all_of(c("log_price", rf_features))) %>%
  drop_na()

rf_df$neighborhood <- as.factor(rf_df$neighborhood)

set.seed(42)
train_index <- createDataPartition(rf_df$log_price, p = 0.8, list = FALSE)
train_data <- rf_df[train_index, ]
test_data <- rf_df[-train_index, ]

rf_model <- randomForest(
  log_price ~ .,
  data = train_data,
  ntree = 150,
  importance = TRUE
)

rf_pred_log <- predict(rf_model, newdata = test_data)
rf_pred_nis <- exp(rf_pred_log)
rf_actual_nis <- exp(test_data$log_price)

rf_mae <- mean(abs(rf_actual_nis - rf_pred_nis))
rf_rmse <- sqrt(mean((rf_actual_nis - rf_pred_nis)^2))
rf_mape <- mean(abs((rf_actual_nis - rf_pred_nis) / rf_actual_nis)) * 100
rf_r2 <- cor(rf_actual_nis, rf_pred_nis)^2

cat("--- Random Forest Evaluation ---\n")
cat(sprintf("R-squared: %.3f\n", rf_r2))
cat(sprintf("MAE: %.0f NIS\n", rf_mae))
cat(sprintf("RMSE: %.0f NIS\n", rf_rmse))
cat(sprintf("MAPE: %.2f%%\n", rf_mape))
```

# 19. Random Forest feature importance

```{r , message=FALSE, warning=FALSE, fig.width=10, fig.height=6}

# Extract feature importance directly from the Random Forest model object
importance_matrix <- rf_model$importance

# Choose the importance column
# For regression Random Forest, usually "IncNodePurity" exists
if ("IncNodePurity" %in% colnames(importance_matrix)) {
  importance_col <- "IncNodePurity"
} else {
  importance_col <- colnames(importance_matrix)[1]
}

rf_importance_df <- data.frame(
  Feature = rownames(importance_matrix),
  Importance = importance_matrix[, importance_col]
) %>%
  arrange(desc(Importance))

rf_importance_df$Feature <- factor(
  rf_importance_df$Feature,
  levels = rev(rf_importance_df$Feature)
)

# Generate the styled Feature Importance plot with color gradient
ggplot(rf_importance_df, aes(x = Importance, y = Feature, fill = Importance)) +
  geom_col(alpha = 0.85, color = "black", width = 0.7) +
  scale_fill_gradient(low = "#ADD8E6", high = "#8B0000") +
  scale_y_discrete(labels = function(x) gsub("log_walking_distance_", "log_", x)) +
  labs(
    title = "Feature Importance in Random Forest Model",
    subtitle = paste("Based on", importance_col, "Metric"),
    x = paste("Importance (", importance_col, ")", sep = ""),
    y = "Feature",
    fill = "Importance Level"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 22, hjust = 0.5),
    plot.subtitle = element_text(size = 16, hjust = 0.5),
    axis.title.x = element_text(size = 16, face = "bold"),
    axis.title.y = element_text(size = 16, face = "bold"),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 13),
    legend.position = "right"
  )

rf_importance_df
```

# 20. XGBoost model

```{r , message=FALSE, warning=FALSE}
xgb_df <- rf_df

set.seed(42)
train_index <- createDataPartition(xgb_df$log_price, p = 0.8, list = FALSE)
xgb_train <- xgb_df[train_index, ]
xgb_test  <- xgb_df[-train_index, ]

# Convert categorical variables to dummy variables
x_train <- model.matrix(log_price ~ . - 1, data = xgb_train)
y_train <- xgb_train$log_price

x_test <- model.matrix(log_price ~ . - 1, data = xgb_test)
y_test <- xgb_test$log_price

# Make sure test matrix has the same columns as train matrix
missing_cols <- setdiff(colnames(x_train), colnames(x_test))
for (col in missing_cols) {
  x_test <- cbind(x_test, 0)
  colnames(x_test)[ncol(x_test)] <- col
}

extra_cols <- setdiff(colnames(x_test), colnames(x_train))
if (length(extra_cols) > 0) {
  x_test <- x_test[, !(colnames(x_test) %in% extra_cols), drop = FALSE]
}

x_test <- x_test[, colnames(x_train), drop = FALSE]

# Create XGBoost matrices
xgb_train_matrix <- xgb.DMatrix(data = x_train, label = y_train)
xgb_test_matrix  <- xgb.DMatrix(data = x_test, label = y_test)

# Train XGBoost model
xgb_params <- list(
  objective = "reg:squarederror",
  eta = 0.05,
  max_depth = 3,
  subsample = 0.8,
  colsample_bytree = 0.8
)

xgb_model <- xgb.train(
  params = xgb_params,
  data = xgb_train_matrix,
  nrounds = 100,
  verbose = 0
)

# Predictions
xgb_pred_log <- predict(xgb_model, newdata = xgb_test_matrix)

# Convert log price back to NIS
xgb_pred_nis <- exp(xgb_pred_log)
xgb_actual_nis <- exp(y_test)

# Evaluation metrics
xgb_mae <- mean(abs(xgb_actual_nis - xgb_pred_nis))
xgb_rmse <- sqrt(mean((xgb_actual_nis - xgb_pred_nis)^2))
xgb_mape <- mean(abs((xgb_actual_nis - xgb_pred_nis) / xgb_actual_nis)) * 100
xgb_r2 <- cor(xgb_actual_nis, xgb_pred_nis)^2

cat("--- XGBoost Evaluation ---\n")
cat(sprintf("R-squared: %.3f\n", xgb_r2))
cat(sprintf("MAE: %.0f NIS\n", xgb_mae))
cat(sprintf("RMSE: %.0f NIS\n", xgb_rmse))
cat(sprintf("MAPE: %.2f%%\n", xgb_mape))
```

# 21. XGBoost feature importance

```{r , message=FALSE, warning=FALSE, fig.width=10, fig.height=6}

# Make sure the XGBoost model exists
if (!exists("xgb_model")) {
  stop("xgb_model does not exist. Run the XGBoost model block before running this block.")
}

# Extract feature importance from XGBoost model
xgb_importance <- xgb.importance(
  feature_names = colnames(x_train),
  model = xgb_model
)

# Plot top 15 important features
xgb_importance_df <- as.data.frame(xgb_importance) %>%
  arrange(desc(Gain)) %>% # או למיין לפי Gain / Frequency בהתאם למה שחישבת
  head(15)

ggplot(xgb_importance_df, aes(x = Gain, y = reorder(Feature, Gain), fill = Gain)) +
  geom_col(alpha = 0.85, color = "black", width = 0.7) +
  scale_fill_gradient(low = "#ADD8E6", high = "#8B0000") +
  scale_y_discrete(labels = function(x) gsub("log_walking_distance_", "log_", x)) +
  labs(
    title = "Feature Importance in XGBoost Model",
    subtitle = "Based on Gain Metric",
    x = "Importance (Gain)",
    y = "Feature",
    fill = "Importance Level"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 22, hjust = 0.5),
    plot.subtitle = element_text(size = 16, hjust = 0.5),
    axis.title.x = element_text(size = 16, face = "bold"),
    axis.title.y = element_text(size = 16, face = "bold"),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 13),
    legend.position = "right"
  )

# Show importance table
xgb_importance
```

# 22.Tree Model comparison

```{r , message=FALSE, warning=FALSE}
# Check that Random Forest metrics exist
if (!exists("rf_r2") | !exists("rf_mae") | !exists("rf_rmse") | !exists("rf_mape")) {
  stop("Random Forest metrics are missing. Run the Random Forest evaluation block first.")
}

# Check that XGBoost metrics exist
if (!exists("xgb_r2") | !exists("xgb_mae") | !exists("xgb_rmse") | !exists("xgb_mape")) {
  stop("XGBoost metrics are missing. Run the XGBoost evaluation block first.")
}

model_comparison <- data.frame(
  Model = c("Random Forest", "XGBoost"),
  R_squared = c(rf_r2, xgb_r2),
  MAE_NIS = c(rf_mae, xgb_mae),
  RMSE_NIS = c(rf_rmse, xgb_rmse),
  MAPE_Percent = c(rf_mape, xgb_mape)
)

model_comparison
```
